---
title: "Lab_03"
author: "Fantastic Four"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(tidyverse)
library(ggplot2)
source("http://www.openintro.org/stat/data/cdc.R")
#cdc <- as_tibble(cdc)
```

# guiding question

In this report we want to find the relationship between people who have smoked at least 100 cigarettes in their life and how it can affect their general health, their health plan coverage, their weight, and frequency of exercise. This report will also reveal the likelihood of people of different ages to have smoked at least 100 cigarettes.

# answer to the guiding question
including the plot

# Recommendation
letter to the CDC, or argument of similar nature

# Individual Findings


## Lexie Marinelli
```{r}
# someone can take this part, I decided to adjust a little: ggplot(data= cdc, mapping =aes(x=age, y=hlthplan)) +geom_smooth()+ facet_wrap(~smoke100)

ggplot(data= cdc)+ stat_count(mapping= aes(x=hlthplan)) + facet_grid(~smoke100) +geom_bar(mapping= aes(x=hlthplan, fill= genhlth), position= "dodge") +theme(aspect.ratio=1) + ggtitle(label = "             Smokers vs. Nonsmokers",subtitle= "        Group 0= Nonsmokers        |          Group 1= Smokers") +xlab(label= "Health Plan") +ylab(label= "Amount of people")
```
  
  This graph portrays that nonsmokers have a higher amount of people in excellent and very good health as well as having a lower amount of people with fair and poor health. This graph also shows that the majority of people, in general, have health insurance not depending on whether they smoke or not. Smokers do have a lower amount of people who have a health plan, but not by a significant amount. The main difference between the smokers and nonsmokers is the general health distribution among the individuals, nonsmokers having a higher amount in the upper quadrant and a lower amount of people in the lower quadrant compared to people who have smoked at least 100 cigarettes.
  
## Lindsay Gettel

## Zhenlong Li
I made a barplot using amount of people versus weight by divding into smokers and non smokers and a barplot using amount of people versus weight by divding into exercisers and non exercisers.

I found that people who smoked at least 100 cigars above good health condition are less than the people who did not. What's more, for condition below good, the amount of people who smoked are larger than people who did not. 

In terms of weight, I found that people who weighed around 150 pounds without smoking are mojority of the people who have excellent health condition, while people who weighed around 150 pounds smoked at least 100 cigars are majority of the people who have poor health condition.

Additionaly, for people who exercised in the past month, majority of them has the health condtion above good. For people who did not exercised in the past month, majority of them has the health condtion below good.

Thus, we can conclude that smoking and exercise did influence a person's health but weight may not be a strong factor to a person's health.

Hers's a plot:
```{r}
ggplot(cdc, aes(weight, fill = genhlth)) +
  geom_histogram(binwidth = 50) + facet_grid(~smoke100) + ggtitle(label = "Smokers' health and weight compared to nonsmokers ", subtitle = "Created by Li") +xlab(label= "weight") +ylab(label= "Amount of people (out of 20,000)")
```
And another one:
```{r}
ggplot(cdc, aes(weight, fill = genhlth)) +
  geom_histogram(binwidth = 50) + facet_grid(~exerany) + ggtitle(label = "Exercisers' health and weight compared to nonexercisers ", subtitle = "Created by Li") +xlab(label= "weight") +ylab(label= "Amount of people (out of 20,000)")
```

## Scott Baker
The three following graphs are similarly composed but I found them to be visually appealing to me using the jitter function and a color function with another variable.

Here are my general findings, which tend to follow intuition:

* There is a larger density of people who smoke that result in 'poor' health
* There is a smaller density of people who are in 'poor' health and have a health plan
* It is more likely that older people have poor health compared to younger people
* In general, there are more people who weight more in the 'poor' health category

The following graph shows that there is a general increase in people who smoke and their general health exam--more light blue dots compared to dark blue dots where the person's health is considered 'poor':
```{r, fig.width=10, fig.height=3,echo=FALSE}
ggplot(data = cdc, mapping = aes(x = weight,y=genhlth,color=smoke100)) + geom_jitter() + ggtitle(label = "Comparing weight to general health with visualization of smoking") +xlab("Weight(lbs)") + ylab("General Health Exam Result")
```

The following plot shows that there is a slight decrease in dark blue dots (no health plan) as the general health exam result gets worse.

```{r, fig.width=10, fig.height=3,echo=FALSE}
ggplot(data = cdc, mapping = aes(x = weight,y=genhlth,color=hlthplan)) + geom_jitter() + ggtitle(label = "Comparing weight to general health with visualization of health plan") +xlab("Weight(lbs)") + ylab("General Health Exam Result")
```
The following graph shows that there is a decrease in health exam result as age goes up--fewer darker dots in the 'poor' catagory than in the 'excellent' catagory.

```{r, fig.width=10,fig.height=3,echo=FALSE}
ggplot(data = cdc, mapping = aes(x = weight,y=genhlth,color=age)) + geom_jitter() + ggtitle(label = "Comparing weight to general health with visualization of age") +xlab("Weight(lbs)") + ylab("General Health Exam Result")

```

# Contributions