---
title: "Lab_14"
author: "Fantastic Four"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(purrr)
options(na.action = na.warn)

COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")

```

# Guiding Question
##What factors lead to on time arrivals? What factors cause flight delays?
It is important to isolate the different events which result in delayed flights and prompt flights in order to improve procedures to make them more effective and give a better assurance that flights will remain on schedule. This information could also help schedule future flights as well as preserve all assets which go into operating airports and flights. The factors we examined were: time, airline, weather, flight time, and length of flight. Although it can be difficult to completely isolate each of these factors (as, in the real world factors are compound), we tried to isolate these elements the best we could. 

# Conclusion
This graph best shows that the main factor of the longest flight delays are caused by weather. The chart below contains the top 10 longest delays in 2017, most of which occur in December through Feburary. This information goes along with what we have found below in the individual findings section about the months relating to amount of delays. There are still points that lie outside of the linear trend which tells us that weather is not the only factor contributing to the delay time.  Furthermore, there seems to be a strong correlation (.957) between the departure delay of a flight and the tardiness at DIA--which means that DIA might not be the problem with late flights.

```{r, echo=FALSE}
modified_data <-COflights %>%
  filter(ARR_DELAY >=15 | DEP_DELAY>=15, MONTH, DAY_OF_MONTH, !WEATHER_DELAY==0)%>%
  mutate(total_delay= ARR_DELAY+DEP_DELAY)

ggplot(data= modified_data, mapping=aes(x= WEATHER_DELAY, y= total_delay)) +geom_point()

modified_data%>%
  filter(total_delay>=2222) %>%
  select(MONTH, DAY_OF_MONTH, total_delay)%>%
  print(MONTH, total_delay)
```

# Overall Findings
There seems to be many factors which all contribute to delayed flights. There are the mesured delays such as weather, security, mechanics, and air traffic. There is also a correlation between the month of the year and the amount of flight delays. January seems to have the most delays, and then the delays decrese until April which then begin to increse until August, followed by incresing delays until December. There is also relevant that the origin and destination of flights impacted the delays, for example, flights from Denver are more likely to be delayed than flights into Denver. There is also a relationship between flight length and how much the flight is delayed. On average two hour flights had the most amount of delays. Additionally, JetBlue and Virgin Airlines had the worst proportion of late flights (~27% and ~25% respectively) into Denver.

#Recommendations
Dear DIA Administraion,
The following observations were made through our analysis:

* Weather is not controlable but is a large factor in flight delays--helps to be more prepared in cases of snow, etc. to make sure you minimize financial loss in operations.
* Effective crowd management could improve your delay time at busy times of the year because there seems to be more late flights at those times (summer and the holidays)
* Solve the domino effect by controlling departure delays at the origin, therefore preventing more arrival delays.
* Examine your contracts (so you do not lose money) with JetBlue, Virgin, and Spirit since they have the highest proportions of late flights.
* Think about examining your contracts with Southwest and United since they have a large mass of late flights.

# Individual Findings

## Lexie Marinelli
```{r, echo=FALSE}
COflights %>%
  filter (MONTH, ARR_DELAY >=15) %>%
  count (MONTH) %>%
  mutate("Number of delayed flights"= n)%>%
  ggplot() + geom_smooth(mapping= aes(x= MONTH, y= n), se=FALSE)+geom_point(mapping=aes(x=MONTH, y=n))+ ggtitle(label="Trend of delayed flights throughout the year") +xlab(label="Month")+ ylab(label= "Number of Delayed Flights")

```

This graph shows the trend of the number of flights delayed in each month. January tops the other months with the number of flights delayed, it drops down to lower values until April. After April there is close to a linear increase in the number of delayed flights going until August, due to summer vacationers. September drops down significantly, most likely it is not a popular month to travel, as well as November since it has the lowest number of delayed flights. December increases in their number of delayed flights due to the hectic holiday season travellers.  

## Lindsay Gettel
How does a late aircraft delay affect the arrival delay of flightes into Denver?
This is a significant question in order to consider how previous flights affect the timing and schedules of others. If there is a relationship between late aircrafts and arrival delays, then precautions can be taken to avoid late aircrafts, and possibly to adjust precedures inbetween one flight and another to help minumize the domino affect of delayed flights.
```{r, echo= FALSE, message=FALSE, warning=FALSE}
DEN<-COflights%>%
  filter(DEST=="DEN", !is.na(ARR_DELAY))
COflights%>%
  filter(DEST=="DEN", !is.na(ARR_DELAY))%>%
  ggplot()+ geom_point(aes(x=LATE_AIRCRAFT_DELAY, y=ARR_DELAY))

linearModel<-lm(ARR_DELAY~LATE_AIRCRAFT_DELAY, data=DEN)

denPred<-DEN%>%
  data_grid(LATE_AIRCRAFT_DELAY)%>%
  add_predictions(linearModel)

denRes<-DEN%>%
  add_residuals(linearModel)

ggplot(DEN, aes(LATE_AIRCRAFT_DELAY))+ geom_point(aes(y=ARR_DELAY))+ geom_line(aes(y=pred), data=denPred, color="blue", size=1)+ ggtitle("Arrival delay and delayed aircraft with liner model projection")

```
The above graph illustrates the linear model of arrival delay regressed on late aircraft delay. The linear fit seams to model a portion of the data, but is not ideal, it demonsrates a slight positive relationship between late aircraft and arrival delays. The model does not fit due to the vertical spread of plots at which there was no delay due to late aicrafts, but there was still a significant arrival delay This indicates that increased late aircraft delays cause arrival delays, however, it is not the singular factor in late arrivals. 
Therefore, on time aircrafts, help minimize late arrivals and increase the chance a flight arrives ontime in Denver.

## Zhenlong Li
Here's a plot showing that most delayed flights occured where flying distance is between 750 mile to 1000 miles:
```{r, echo=FALSE}
filter(COflights,ARR_DELAY >=15)%>%
  select(DISTANCE)%>%
  count (DISTANCE) %>%
  ggplot()+ geom_jitter(mapping = aes(x= DISTANCE, y = n)) + geom_smooth(mapping = aes(x= DISTANCE, y= n), method = 'loess', se = FALSE) + ylab(label= "Number of Delayed Flights")
dis<-COflights%>% filter(DISTANCE>=750 & DISTANCE<=1000)%>% count()
all<-COflights%>% count()
p1 = dis / all
p1
```
And the probability of that flying distance between 750 mile to 1000 miles will have 35% to be delayed.

Here's a plot showing that the actual air time greater than 300 mins will have less on-time flights:
```{r, echo=FALSE}
filter(COflights,ARR_DELAY <= 14)%>%
  select(ACTUAL_ELAPSED_TIME)%>%
  count (ACTUAL_ELAPSED_TIME) %>%
  ggplot()+ geom_col( mapping = aes(x= ACTUAL_ELAPSED_TIME, y= n), width = 1)+ geom_smooth(mapping = aes(x= ACTUAL_ELAPSED_TIME, y= n), method = 'loess', se = FALSE) +xlab(label="Elapsed Time")+ ylab(label= "Number of On-time Flights")
time<-COflights%>% filter(ACTUAL_ELAPSED_TIME>=300)%>% count()
all<-COflights%>% count()
p2 = time / all
p2
```
And the probability of that the actual air time greater than 300 mins will have 0.65% to be on-time.


## Scott Baker

First, I take a look at the number of flights that arrived late (more than 15 minutes) into Denver based on airline. Here is the graph:

```{r,echo=FALSE}
stuff <- filter(COflights,ARR_DELAY >= 15, DEST == "DEN") %>%
  select(CARRIER)
ggplot(stuff,mapping = aes(x = CARRIER)) + geom_bar() + geom_text(stat='count',aes(label=..count..),vjust = -0.5) + ggtitle("Number of ARR_DELAY >= 15 flights into Denver per Airline")
```

The first major observation is that airlines "WN", "UA", and "OO" have the most flights delayed landing.
* Southwest (WN): 12577
* United (UA): 6889
* SkyWest (OO): 6070

Although this confirms personal experience that some United and Southwest flights I've been on have been delayed, this does not provide any insight into the probabilities that a certain airline will have a delayed arrival. Furthermore, let us look at the actual proportion that a certain airline's flight to Denver is late. Observe the following table:

```{r,echo=FALSE}
allflights <- filter(COflights,DEST == "DEN") %>%
  select(CARRIER)

tb1 <- allflights %>% table()
tb2 <- stuff %>% table()
tb2/tb1

```

Here, we can see that "B6", "VX", and "NK" have the highest rate of late flights.
* JetBlue (B6): 27.31%
* Virgin (VX): 25.36%
* Spirit (NK): 20.24%

From this, we find that actually JetBlue, Virgin, and Spirit have the most delayed arrivals with respect to total flights--NOT United, Southwest, and SkyWest.

The second part of my analysis is to show that there is a strong correlation between the departure delay and the tardiness of their arrival:

```{r,echo=FALSE}
arrdepdelay <- filter(COflights,DEST=="DEN"&!is.na(ARR_DELAY)&!is.na(DEP_DELAY)) %>% select(ARR_DELAY,DEP_DELAY)

ggplot(arrdepdelay,mapping=aes(x=ARR_DELAY,y=DEP_DELAY)) + geom_point()+ ggtitle("Arrival Delay vs. Departure Delay")
cor(arrdepdelay$ARR_DELAY,arrdepdelay$DEP_DELAY)
```

From this we have two results. First, the correlation value between arrival delay (in Denver) and departure delay is .957, which is a strong correlation. Second, the implication of this is that there sometimes might not be anything DIA can do better to ensure more flights are on time--it could all be a product of the departure airport and their inability to get planes off in time. Although, this observation is generic and more specifics would be needed to confirm.

# Contributions
*Lindsay: Created a model to add predictions and determine if there is linear relationship between arrival delays to Denver and aircraft delays. 

* Lexie: Created individual plot and decription of graph. Also helped create the conclusion plot with help conclusion paragraph and recommendation ideas.

* Li: Creted two plots to see how distance and actual elapsed time affected whether a flight is delayed or on-time. Checking my findings by using calcluating the probability of each event.

* Scott: Besides the normal organization of the git and knitting/submitting the final file, I took care of the recommendation to DIA officials. I also examined airline-by-airline arrival delay.
