---
title: "Lab_13"
author: "Fantastic Four"
output: html_document
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(forcats)
library(stringr)
library(magrittr)

#plain gapminder data
gapminder <- gapminder::gapminder_unfiltered

#scott's food gapminder data
gmFood <- as_data_frame(read_csv("https://raw.githubusercontent.com/bakerscott/CU-APPM2720/master/m5/Lab13/Gapminder_Food_Consumption.csv"))

#lexie's data
gdp<- read_csv("https://raw.githubusercontent.com/bakerscott/CU-APPM2720/master/m5/Lab13/gdp_total_ppp.csv")
hiv<- read_csv("https://raw.githubusercontent.com/bakerscott/CU-APPM2720/master/m5/Lab13/hiv_plwha.csv")

#lindsay's data
femaleBMI<-read.csv("https://raw.githubusercontent.com/bakerscott/CU-APPM2720/master/m5/Lab13/Indicator_BMI_female.csv", header=TRUE, check.names=FALSE)
maleBMI<-read.csv("https://raw.githubusercontent.com/bakerscott/CU-APPM2720/master/m5/Lab13/Indicator_BMI_male.csv", header = TRUE, check.names = FALSE)
```

# How can CU's involvements help improve living around the world?

```{r}
#this section will be for giving reason to improving lives

```





# Subquestions

## How has food consumption changed in the last forty years? (Scott)

Has there been an increase in food consumption from the 1960's to the 2000's in the United States? This question is important in assessing the overall health of students at CU and people in our country over the last forty years. By looking at this question, potential improvments could be made at CU with regards to food consumption and health. Additionally, answering this question may give logic to funding food sustainability research.

The null hypothesis we wish to test is "there is no change in food consumption from 1960 to 2000 in the United States".

The test statistic is ten "unlabeled" values for food consumption over the last forty years. Here is the histogram of 1000 iterations of the test statistic.

```{r,warning=FALSE,msg=FALSE,fig.width=10,fig.height=4}
#this is for cleanup:
gmFood <- as.data.frame(gmFood)
#united states food consumption
uS <- gmFood[241,] #snag the row US
uS <- uS[-1] #get rid of the "United States" label

dfUS <- data_frame(year=integer(),
                    energy=double()
                   )
tempY <- 1960
for(i in seq_along(uS)){
  dfUS[i,2] <- uS[1,i]
  dfUS[i,1] <- tempY + i
}
#data for the 60's:
uS60 <- filter(dfUS, (year < 1970 & year >= 1960))
#data for the 2k's:
uS2K <- filter(dfUS, (year >= 2000))
#data for "fake data":
dfUS <- filter(dfUS, (year >= 2000) || (year < 1970 & year >= 1960))
i <- 0
#this is the number of iterations:
len <- 1000 
output <- as.data.frame(1:len)
#output
colnames(output) <- c("values")
#ITERATION:
while(i < len){
  temp <- sample_n(dfUS,10,replace=TRUE)
  meanT <- mean(temp$energy)
  output[i+1,1] <- meanT
  i <- i + 1
}
#output
#percentiles setup:
m60 <- mean(uS60$energy)
m2k <- mean(uS2K$energy)
pm60 <- 0
pm2k <- 0
j <- 0
#percentile checking
while(j < len){
  if(output[j+1,1] < m60){
    pm60 <- pm60 + 1
  }
  else if(output[j+1,1] < m2k){
    pm2k <- pm2k + 1
  }
  j <- j + 1
}
#normalization to a percentage
pm60 <- pm60 / len*100
pm2k <- pm2k / len*100
#percentile the mean of the 60's life expectancy is:
pm60
#percentile the mean of the 2k's life expectancy is:
pm2k
#plot
ggplot(data=output) + geom_histogram(mapping=aes(values),color="green", bins=30) + 
  geom_vline(aes(xintercept = m60),color="red",lty=2) + 
  geom_vline(aes(xintercept = m2k),color="blue",lty=2) +
  geom_text(aes(x=3700, label="2000's Mean", y=100), colour="blue", angle=90, text=element_text(size=20)) +
  geom_text(aes(x=2900, label="1960's Mean", y=100), colour="red", angle=90, text=element_text(size=20)) +
  xlab("Mean Energy Consumption (kilocals/person per day)") +
  ylab("Count") +
  ggtitle("Sampling of Food Consumption in the US from the 1960's to the 2000's")

#mean of the test stat 1K
MM <- sapply(output,mean)
#standard deviation of the test stat 1K
SS <- sapply(output,sd)

#zscore 60's
z60 <- (m60 - MM)/SS
#ascore 2k's
z2k <- (m2k - MM)/SS
```

The distribution of the test statistic is fairly normal with mean `r MM` and standard deviation `r SS`. It is seen that the mean value for the 1960's is `r m60` and sits at the 0th percentile. This means that there are no values in the distribution of the test statistic that is lower than this mean. Furthermore, the z-score for an assumed normal distribution is `r z60` which tells us that there is a significant difference between average for the 60's and testing average. Additionally, the same analysis is true for the value in the 2000's. The average food consumption in the 2000's is `r m2k` which sits at the 100th percentile (technically 99.9th) and has a z-score of `r z2k`. By assessing these scores, the null hypothesis can be rejected and it can be said that there is a significant difference between food consumption in the 1960's and in the 2000's.

With regard to CU, this increase in consumption on average means two things for our school. First, the school must make sure they are promoting health programs in order to keep students active and not let them consume more food than they can maintain. Second, since there has been an increase in consumption, the school must keep funding research about food sustainability. With a growing population and growing consumption, research on sustatinable food would be a good investment for the future of society.


##Lindsay
```{r}
fBMI<-gather(femaleBMI, -`Country`, key=year, value="femaleBMI")
BMI_f<-rename(fBMI, "country"="Country")
mBMI<-gather(maleBMI, -`Country`, key=year, value="maleBMI")
BMI_m<-rename(mBMI, "country"="Country")
addFBMI<-merge(gapminder, BMI_f, by=c("country"="country", "year"="year"))
BMI.data<-merge(addFBMI, BMI_m, by=c("country"="country", "year"="year"))
```
1. Is there a significantly significant difference between the difference in mean values of the differences in African male and female BMI by from 1987 to 2007?
This is significant in modeling BMI within a population in order to determine possible trends over a given range of time. If there is an increase or decrease in BMI, it would suggest possible harmful changes in population health. If there is no significant statistical difference, then BMI is consistant overtime.
2. There should not be a difference in the mean difference BMI for the countries, the year should not affect the difference which should be zero.
3. The summary of the data wll be the mean of the difference in BMI for each African countries' male and female BMI.
4. After determining the means of the countries BMI difference, drawn repeatidly, the values will be placed in a histogram , and the mean of the distribution shoud occur at zero.
5. For female BMI 40% of the mean difference values are between -0.119 and 0.110, while 40% of male BMI fall between -0.73 and 0.069. For both sexes, almost half the data lies in those ranges, which are very small, which suggests that most of the differences are zero, or very close to zero.
6. The two graphs illustrate the mean value of the randomly selected differences between BMI for males and females from 1987 to 2007. The female BMI is representative of a standard normal distribution, with the mean at 0, this is the expected result of the null hypothisis. This data validates the null hypothsis, that BMI is not dependent on country or year, beacue the randomly sampled data has a mean diffrence in BMI of zero when the labels are stripped off the data.
```{r, echo= FALSE, message=FALSE, warning=FALSE}
set.seed(30)
female.diffBMI<-tibble(country=unique(filter(BMI.data, continent=="Africa", year==1987)$country), one=0, two=0, three=0, four=0)
BSIsample<-vector("double", length=length(female.diffBMI$country))
female.medians<-vector("double", length=1000)
mixBMI<-function(place, yr=c(1987, 2007)){
  difference<-filter(BMI.data, country==place, year==yr[1])%>%select(femaleBMI)-filter(BMI.data, country==place, year==yr[2])%>%select(femaleBMI)
  diff<-as.numeric(difference)
  c(0, diff, -diff, 0)
}
for(j in seq_along(female.diffBMI$country)){
  female.diffBMI[j, 2:5]<-mixBMI(female.diffBMI$country[j])
}
getSample<-function(data, vector){
  for(k in seq_along(vector)){
    vector[k]<-as.numeric(sample(data[sample(nrow(data),1),2:5], 1))
      #sample(female.diffBMI[,2:5],1)
  }
  return(vector)
}
for(n in seq_along(female.medians)){ 
  BSIsample<-getSample(female.diffBMI, BSIsample)
  female.medians[n]<-mean(BSIsample)
}
ggplot()+ geom_histogram(aes(female.medians), bins=70)+ geom_vline(aes(xintercept=0), color="red")+ ggtitle("Means of country female BSI from 1987 to 2007")
quantile(female.medians, probs=c(.3,.7))
```
The above data illustrates that over time there as been little to no change in the Afrian populatin BMI. This suggests that the population has maintained a steady average BMI from 1987-2007, suggesting the overall health of the population has not changed. Using this model to represent a larger demographic, no change in BMI indicates a stability for population health, however, if recalculated and the BMI fuluctuates, it would indicate possible heath issues such as obesity, or malnutrition. To make this model more accurate, it would be benificial to determine the change in BMI for more recent dates and other countries.

##Lexie


```{r, echo= FALSE}

```

Question: How is GDP related to adults with HIV? This is important because it can affect how we address AIDS in developing countries compared to how we treat it in developed countries. 
Null Hypothesis: GDP is not related to the amount of adults with HIV.

```{r, echo= FALSE}
colnames(gdp)[1]<- "country"

gdp <- gdp %>%
  gather(-country, key= "year", value = "totalGDP")

gdp<-gdp %>%
  select(country, year, totalGDP) %>%
  filter(!is.na(totalGDP), year>=1990)

colnames(hiv)[1]<- "country"

hiv<- hiv %>%
  gather(-country, key= "year", value = "HIVcases")

hiv<- hiv %>%
  filter(!is.na(HIVcases), year>= 1990) 
hiv[3] <- transform(as.integer(hiv$HIVcases))
hiv[2]<- transform(as.integer(hiv$year))
hiv[1] <- transform(as.factor(hiv$country))


hivGap <- full_join(gapminder, hiv, by= "country")
hivGap<- hivGap %>%  
  filter(year.x == year.y) %>%
  select(-year.y)

colnames(hivGap)[3]<- "year"

hivgdp <- full_join(gdp, hiv, by= "country")
hivgdp<- hivgdp %>%  
  filter(year.x == year.y) %>%
  select(-year.y)

colnames(hivgdp)[2]<- "year"

hivgdp1 <- hivgdp %>%
  filter(country== "Afghanistan"| country== "Australia" | country== "Brazil" | country == "Congo, Rep" | country == "Cote d'Ivoire" | country=="France" | country== "Haiti" | country== "India" | country=="Kenya" | country== "Nigeria" | country== "Rwanda" | country== "South Africa" | country== "Thailand" | country=="Tanzania" | country=="Switzerland" | country=="United Kingdom" | country== "United States" |country == "Uganda" | country== "Zimbabwe")

gdp1 <- gdp %>%
  filter(country== "Afghanistan"| country== "Australia" | country== "Brazil" | country == "Congo, Rep" | country == "Cote d'Ivoire" | country=="France" | country== "Haiti" | country== "India" | country=="Kenya" | country== "Nigeria" | country== "Rwanda" | country== "South Africa" | country== "Thailand" | country=="Tanzania" | country=="Switzerland" | country=="United Kingdom" | country== "United States" |country == "Uganda" | country== "Zimbabwe")
```

```{r, echo=FALSE,fig.width=5,fig.height=5}

#ggplot(hivgdp1, aes(year, HIVcases, color= country)) + geom_point(alpha=1/3) + geom_smooth(data=gdp1, aes(year, totalGDP, color= country)) + geom_text(show.legend = FALSE, label= countries, nudge_y = 0.1, nudge_x = 0.05)

hivgdpAVG <- hivgdp1 %>%
  group_by(year) %>%
  summarise(avgHIV= mean(HIVcases, na.rm=TRUE), avgGDP= mean(totalGDP, na.rm=TRUE))

totalAvgHivGdp<- hivgdpAVG %>%
  summarise(totalAvgHIV= mean(avgHIV), totalAvgGdp= mean(avgGDP))


stdev<- hivgdpAVG %>%
  summarise(sd(avgHIV), sd(avgGDP))


hivgdpAVG1 <- hivgdpAVG %>%
  mutate(stdHIV= (avgHIV- totalAvgHivGdp[[1]])/ stdev[[1]], stdGDP= (avgGDP - totalAvgHivGdp[[2]])/ stdev[[2]])

countries<- c("Afghanistan", "Australia", "Brazil", "Cote d'Ivoire", "France", "Haiti", "India", "Kenya", "Nigeria", "Rwanda", "South Africa", "Thailand", "Tanzania", "Switzerland", "United Kingdom", "United States", "Uganda", "Zimbabwe")
years<- c("1990","1991","1992", "1993", "1994", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011")

hivgdpAVG1 %>%ggplot(aes(stdGDP, stdHIV)) + geom_point(aes(color= year), size=3) + geom_line() +geom_text(show.legend = FALSE, label= years, nudge_y = 0.1, nudge_x = 0.05)+ xlab("Standardized GDP")+  ylab("Standardized HIV") +ggtitle("GDP vs. HIV in adult population")
```

Test Statistic: Using the standardized values for a selected number of countries GDP and adults with HIV, the graph shows for some time between 1990-2011, the standardized number for GDP and HIV show a linear relationship. As time gets closer to 2000, the trend plateaus and remains slowly increasing for the rest of the plot. 


```{r, echo= FALSE, message=FALSE, warning=FALSE}
#for defining functions (?):
showLevels <- function(data_input){
  return(unique(data_input))
}

summaryStats <- function(data_input,continent_in,start_year,end_year,column_in){
  #main function to return the summary stats for a certain dataset and column for years
  x2 <- filter(data_input,continent == continent_in,year>=start_year,year<=end_year)
  x <- sample_n(x2,10,replace=TRUE)
  df <- data.frame(continent_in,start_year,end_year,mean=mean(x[[column_in]]),median=median(x[[column_in]]),min=min(x[[column_in]]),max=max(x[[column_in]]),IQR=IQR(x[[column_in]]))
  return(df)
}
calc_summaryStats <- function(data_input,continent_in,start_year,end_year,column_in){
  yearA <- start_year
  output_summary <- 0

  while(yearA < end_year){
    output_summary <- rbind(output_summary, summaryStats(data_input,continent_in,yearA,yearA+9,column_in))
    yearA <- yearA + 10
  }
  output_summary <- output_summary[-1,]
  colnames(output_summary) <- c("continent","start_year","end_year","mean","median","minimum","maximum","IQR")
  return(output_summary)
}

runAll_calc_summaryStats <- function(data_input,start_year,end_year,column_in){
  #cont = c("Africa","Asia","Americas","Europe","Oceania")
  output_dataframe <- 0
  fun <- showLevels(data_input$continent)
  for(item in fun){
    output_dataframe <- rbind(output_dataframe, calc_summaryStats(hivGap,item,start_year,end_year,column_in))
  }
  return(output_dataframe)
}

i <- 0
temp <- 0
#while(i < 100){
#  temp <- rbind(temp,calc_summaryStats(hivGap,"Africa",1990,2012,"HIVcases"))
#  temp <- temp[-1,]
#  i <- i + 1
#}

#ggplot(data=temp) + geom_histogram(mapping=aes(mean), bins=20) +facet_grid(~start_year) + geom_vline(aes(xintercept=mean(mean)),color='Red') + ggtitle("Life Expectancy in Africa over time") + ylab("Mean Life Expectancy") + xlab("Year (start of decade)")

hivGapAVG <- hivGap%>%
  group_by(year) %>%
  summarise(avgHIV= mean(HIVcases, na.rm=TRUE), avgGDP= mean(gdpPercap, na.rm=TRUE))
hivGapAVG1<- hivGapAVG %>%
  summarise(totalAvgHIV= mean(avgHIV), totalAvgGDP= mean(avgGDP))



#back-up plan
ggplot(data= hivGapAVG) +geom_histogram(aes(avgHIV), bins=20) +geom_vline(aes(xintercept= hivGapAVG1[[1]]), color= "red") + ggtitle("Average number of adults with HIV over time")
```

In this histogram, the distribution is mostly on the left side of the average, but still close to the median. This calculation is overall a 
accurate representation of the average number of adults with HIV over time. 

##Li

The question is "Did the students (15-24 years old) have a better chance to get employed in the United States from the 1990's to the 2000's?"

The Null Hypothesis is "There is no change between the employment rate for young people in US between the 1990's and the 2000's"

The test statistic of the data is a mean of a sample of 10 values from the data.

```{r,warning=FALSE,message=FALSE}
employment <- read_csv("https://github.com/bakerscott/CU-APPM2720/raw/master/m5/Lab13/indicator_t%2015-24%20employ.csv")

colnames(employment) <- c("country","1991","1992","1993","1994", "1995","1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007")

rate <- gather(employment, -'country', key = "year", value = "EmploymentRate")
rate$year <- as.numeric(rate$year)
rate$EmploymentRate <- as.numeric(rate$EmploymentRate)

Data <- filter(rate, country == "United States", (year < 2000 & year >= 1900) | (year < 2100 & year >= 2000))

Data90 <- filter(rate, country == "United States", (year < 2000 & year >= 1900))

Data00 <- filter(rate, country == "United States", (year < 2100 & year >= 2000))

i <- 0
j <- 0
n <- 1000

output <- as.data.frame(1:n)
colnames(output) <- c("values")

while(i < n){
  temp <- sample_n(Data,10,replace=TRUE)
  meanT <- mean(temp$EmploymentRate)
  output[i+1,1] <- meanT
  i <- i + 1
}


m90 <- mean(Data90$EmploymentRate)
m00 <- mean(Data00$EmploymentRate)
pm90 <- 0
pm00 <- 0


while(j < n){
  if(output[j+1,1] < m90){
    pm90 <- pm90 + 1
  }
  else if(output[j+1,1] < m00){
    pm00 <- pm00 + 1
  }
  j <- j + 1
}


pm90 <- pm90 / n*100
pm00 <- pm00 / n*100

pm90

pm00



ggplot(data=output) + geom_histogram(mapping=aes(values), bins=20) + 
  geom_vline(aes(xintercept = m90),color="red",lty=2) + 
  geom_vline(aes(xintercept = m00),color="blue",lty=2) +
  geom_text(aes(x=m00, label="2000", y=100), colour="blue", angle=90, text=element_text(size=16)) +
  geom_text(aes(x=m90, label="1990", y=100), colour="red", angle=90, text=element_text(size=16)) +
  xlab("Mean Employment Rate") +
  ylab("Count") +
  ggtitle("Sampling of Employment Rate in US from the 1990's to the 2000's")
  
```

From this above distribution, the 1990's employment rate was at `r m90` where the percentile was `r pm90`% and this means that there are no values in the distribution of the test statistic that is greater than this mean. The 2000's employment rate was at `r m00` where the percentile was `r pm00`% and this means that there are no values in the distribution of the test statistic that is smaller than this mean. Therefore we reject the null hypothesis and conclude that the employment rate for students in US is decresing and the employment situation is not optimistic in recent decade years.

In term of the Chancellor of CU Boulder's request for help him and others around campus "benefit humanity", we would encourage students to go to graduate school after graduating and for the students who would like to work. Also, we would suggest that the career service of CU should be cherished and developed with more investment in the future to help students to have a better chance to get a decent job.

# Contributions
* Lindsay: Created functions to find differences of values, then randomly select data to calculate the mean, then plotted values. Also merged datasets and found percentiles for data.

* Lexie: 

* Li: Used interating functions to testing the hypothesis about whether the employment rate of students between 1991 and 2007 is decreasing or increasing and related it with the overall question about CU.

* Scott:
