---
title: "Lab_13"
author: "Fantastic Four"
output: html_document
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(forcats)
library(stringr)
library(magrittr)

#plain gapminder data
gapminder <- gapminder::gapminder_unfiltered

#scott's food gapminder data
gmFood <- as_data_frame(read_csv("https://raw.githubusercontent.com/bakerscott/CU-APPM2720/master/m5/Lab13/Gapminder_Food_Consumption.csv"))

#lexie's data
gdp<- read_csv("https://raw.githubusercontent.com/bakerscott/CU-APPM2720/master/m5/Lab13/gdp_total_ppp.csv")
hiv<- read_csv("https://raw.githubusercontent.com/bakerscott/CU-APPM2720/master/m5/Lab13/hiv_plwha.csv")
```

# How can CU's involvements help improve living around the world?

```{r}
#this section will be for giving reason to improving lives

```





# Subquestions

## How has food consumption changed in the last fourty years? (Scott)

Has there been an increase in food consumption from the 1960's to the 2000's in the United States? This question is important in assessing the overall health of students at CU over the last fourty years. By looking at this question, potential improvments could be made at CU with regards to food consumption and health.

The null hypothesis we wish to test is "there is no change in food consumption from 1960 to 2000 in the United States".

The test statistic is ten "unlabeled" values for food consumption over the last fourty years. Here is the histogram of 1000 iterations of the test statistic.

```{r,warning=FALSE,msg=FALSE,fig.width=10,fig.height=4}
#this is for cleanup:
gmFood <- as.data.frame(gmFood)
#united states food consumption
uS <- gmFood[241,] #snag the row US
uS <- uS[-1] #get rid of the "United States" label

dfUS <- data_frame(year=integer(),
                    energy=double()
                   )
tempY <- 1960
for(i in seq_along(uS)){
  dfUS[i,2] <- uS[1,i]
  dfUS[i,1] <- tempY + i
}
#data for the 60's:
uS60 <- filter(dfUS, (year < 1970 & year >= 1960))
#data for the 2k's:
uS2K <- filter(dfUS, (year >= 2000))
#data for "fake data":
dfUS <- filter(dfUS, (year >= 2000) || (year < 1970 & year >= 1960))
i <- 0
#this is the number of iterations:
len <- 1000 
output <- as.data.frame(1:len)
#output
colnames(output) <- c("values")
#ITERATION:
while(i < len){
  temp <- sample_n(dfUS,10,replace=TRUE)
  meanT <- mean(temp$energy)
  output[i+1,1] <- meanT
  i <- i + 1
}
#output
#percentiles setup:
m60 <- mean(uS60$energy)
m2k <- mean(uS2K$energy)
pm60 <- 0
pm2k <- 0
j <- 0
#percentile checking
while(j < len){
  if(output[j+1,1] < m60){
    pm60 <- pm60 + 1
  }
  else if(output[j+1,1] < m2k){
    pm2k <- pm2k + 1
  }
  j <- j + 1
}
#normalization to a percentage
pm60 <- pm60 / len*100
pm2k <- pm2k / len*100
#percentile the mean of the 60's life expectancy is:
pm60
#percentile the mean of the 2k's life expectancy is:
pm2k
#plot
ggplot(data=output) + geom_histogram(mapping=aes(values),color="green", bins=30) + 
  geom_vline(aes(xintercept = m60),color="red",lty=2) + 
  geom_vline(aes(xintercept = m2k),color="blue",lty=2) +
  geom_text(aes(x=3700, label="2000's Mean", y=100), colour="blue", angle=90, text=element_text(size=20)) +
  geom_text(aes(x=2900, label="1960's Mean", y=100), colour="red", angle=90, text=element_text(size=20)) +
  xlab("Mean Energy Consumption (kilocals/person per day)") +
  ylab("Count") +
  ggtitle("Sampling of Food Consumption in the US from the 1960's to the 2000's")

#mean of the test stat 1K
MM <- sapply(output,mean)
#standard deviation of the test stat 1K
SS <- sapply(output,sd)

#zscore 60's
z60 <- (m60 - MM)/SS
#ascore 2k's
z2k <- (m2k - MM)/SS
```

The distribution of the test statistic is fairly normal with mean `r MM` and standard deviation `r SS`. It is seen that the mean value for the 1960's is `r m60` and sits at the 0th percentile. This means that there are no values in the distribution of the test statistic that is lower than this mean. Furthermore, the z-score for an assumed normal distribution is `r z60` which tells us that there is a significant difference between average for the 60's and testing average. Additionally, the same analysis is true for the value in the 2000's. The average food consumption in the 2000's is `r m2k` which sits at the 100th percentile (technically 99.9th) and has a z-score of `r z2k`. By assessing these scores, the null hypothesis can be rejected and it can be said that there is a significant difference between food consumption in the 1960's and in the 2000's.

With regard to CU, this increase in consumption on average means two things for our school. First, the school must make sure they are promoting health programs in order to keep students active and not let them consume more food than they can maintain. Second, since there has been an increase in consumption, the school must keep funding research about food sustainability. With a growing population and growing consumption, research on sustatinable food would be a good investment for the future of society.


##Lindsay




##Lexie


```{r, echo= FALSE}

```

Question: How is GDP related to adults with HIV? This is important because it can affect how we address AIDS in developing countries compared to how we treat it in developed countries. 
Null Hypothesis: GDP is not related to the amount of adults with HIV.

```{r, echo= FALSE}
colnames(gdp)[1]<- "country"

gdp <- gdp %>%
  gather(-country, key= "year", value = "totalGDP")

gdp<-gdp %>%
  select(country, year, totalGDP) %>%
  filter(!is.na(totalGDP), year>=1990)

colnames(hiv)[1]<- "country"

hiv<- hiv %>%
  gather(-country, key= "year", value = "HIVcases")

hiv<- hiv %>%
  filter(!is.na(HIVcases), year>= 1990) 
hiv[3] <- transform(as.integer(hiv$HIVcases))
hiv[2]<- transform(as.integer(hiv$year))
hiv[1] <- transform(as.factor(hiv$country))


hivGap <- full_join(gapminder, hiv, by= "country")
hivGap<- hivGap %>%  
  filter(year.x == year.y) %>%
  select(-year.y)

colnames(hivGap)[3]<- "year"

hivgdp <- full_join(gdp, hiv, by= "country")
hivgdp<- hivgdp %>%  
  filter(year.x == year.y) %>%
  select(-year.y)

colnames(hivgdp)[2]<- "year"

hivgdp1 <- hivgdp %>%
  filter(country== "Afghanistan"| country== "Australia" | country== "Brazil" | country == "Congo, Rep" | country == "Cote d'Ivoire" | country=="France" | country== "Haiti" | country== "India" | country=="Kenya" | country== "Nigeria" | country== "Rwanda" | country== "South Africa" | country== "Thailand" | country=="Tanzania" | country=="Switzerland" | country=="United Kingdom" | country== "United States" |country == "Uganda" | country== "Zimbabwe")

gdp1 <- gdp %>%
  filter(country== "Afghanistan"| country== "Australia" | country== "Brazil" | country == "Congo, Rep" | country == "Cote d'Ivoire" | country=="France" | country== "Haiti" | country== "India" | country=="Kenya" | country== "Nigeria" | country== "Rwanda" | country== "South Africa" | country== "Thailand" | country=="Tanzania" | country=="Switzerland" | country=="United Kingdom" | country== "United States" |country == "Uganda" | country== "Zimbabwe")
```

```{r, echo=FALSE,fig.width=5,fig.height=5}

#ggplot(hivgdp1, aes(year, HIVcases, color= country)) + geom_point(alpha=1/3) + geom_smooth(data=gdp1, aes(year, totalGDP, color= country)) + geom_text(show.legend = FALSE, label= countries, nudge_y = 0.1, nudge_x = 0.05)

hivgdpAVG <- hivgdp1 %>%
  group_by(year) %>%
  summarise(avgHIV= mean(HIVcases, na.rm=TRUE), avgGDP= mean(totalGDP, na.rm=TRUE))

totalAvgHivGdp<- hivgdpAVG %>%
  summarise(totalAvgHIV= mean(avgHIV), totalAvgGdp= mean(avgGDP))


stdev<- hivgdpAVG %>%
  summarise(sd(avgHIV), sd(avgGDP))


hivgdpAVG1 <- hivgdpAVG %>%
  mutate(stdHIV= (avgHIV- totalAvgHivGdp[[1]])/ stdev[[1]], stdGDP= (avgGDP - totalAvgHivGdp[[2]])/ stdev[[2]])

countries<- c("Afghanistan", "Australia", "Brazil", "Cote d'Ivoire", "France", "Haiti", "India", "Kenya", "Nigeria", "Rwanda", "South Africa", "Thailand", "Tanzania", "Switzerland", "United Kingdom", "United States", "Uganda", "Zimbabwe")
years<- c("1990","1991","1992", "1993", "1994", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011")

hivgdpAVG1 %>%ggplot(aes(stdGDP, stdHIV)) + geom_point(aes(color= year), size=3) + geom_line() +geom_text(show.legend = FALSE, label= years, nudge_y = 0.1, nudge_x = 0.05)+ xlab("Standardized GDP")+  ylab("Standardized HIV") +ggtitle("GDP vs. HIV in adult population")
```

Test Statistic: Using the standardized values for a selected number of countries GDP and adults with HIV, the graph shows for some time between 1990-2011, the standardized number for GDP and HIV show a linear relationship. As time gets closer to 2000, the trend plateaus and remains slowly increasing for the rest of the plot. 


```{r, echo= FALSE, message=FALSE, warning=FALSE}
#for defining functions (?):
showLevels <- function(data_input){
  return(unique(data_input))
}

summaryStats <- function(data_input,continent_in,start_year,end_year,column_in){
  #main function to return the summary stats for a certain dataset and column for years
  x2 <- filter(data_input,continent == continent_in,year>=start_year,year<=end_year)
  x <- sample_n(x2,10,replace=TRUE)
  df <- data.frame(continent_in,start_year,end_year,mean=mean(x[[column_in]]),median=median(x[[column_in]]),min=min(x[[column_in]]),max=max(x[[column_in]]),IQR=IQR(x[[column_in]]))
  return(df)
}
calc_summaryStats <- function(data_input,continent_in,start_year,end_year,column_in){
  yearA <- start_year
  output_summary <- 0

  while(yearA < end_year){
    output_summary <- rbind(output_summary, summaryStats(data_input,continent_in,yearA,yearA+9,column_in))
    yearA <- yearA + 10
  }
  output_summary <- output_summary[-1,]
  colnames(output_summary) <- c("continent","start_year","end_year","mean","median","minimum","maximum","IQR")
  return(output_summary)
}

runAll_calc_summaryStats <- function(data_input,start_year,end_year,column_in){
  #cont = c("Africa","Asia","Americas","Europe","Oceania")
  output_dataframe <- 0
  fun <- showLevels(data_input$continent)
  for(item in fun){
    output_dataframe <- rbind(output_dataframe, calc_summaryStats(hivGap,item,start_year,end_year,column_in))
  }
  return(output_dataframe)
}

i <- 0
temp <- 0
#while(i < 100){
#  temp <- rbind(temp,calc_summaryStats(hivGap,"Africa",1990,2012,"HIVcases"))
#  temp <- temp[-1,]
#  i <- i + 1
#}

#ggplot(data=temp) + geom_histogram(mapping=aes(mean), bins=20) +facet_grid(~start_year) + geom_vline(aes(xintercept=mean(mean)),color='Red') + ggtitle("Life Expectancy in Africa over time") + ylab("Mean Life Expectancy") + xlab("Year (start of decade)")

hivGapAVG <- hivGap%>%
  group_by(year) %>%
  summarise(avgHIV= mean(HIVcases, na.rm=TRUE), avgGDP= mean(gdpPercap, na.rm=TRUE))
hivGapAVG1<- hivGapAVG %>%
  summarise(totalAvgHIV= mean(avgHIV), totalAvgGDP= mean(avgGDP))



#back-up plan
ggplot(data= hivGapAVG) +geom_histogram(aes(avgHIV), bins=20) +geom_vline(aes(xintercept= hivGapAVG1[[1]]), color= "red") + ggtitle("Average number of adults with HIV over time")
```

In this histogram, the distribution is mostly on the left side of the average, but still close to the median. This calculation is overall a 
accurate representation of the average number of adults with HIV over time. 

##Li


# Contributions
* Lindsay: 

* Lexie: 

* Li: 

* Scott:
