---
title: "Lab_05"
author: "Fantastic Four"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#taken straight from Lab_5.html provided:
rm(list=ls()) # clears your environment variables
cat('\014')   # clears the console
suppressPackageStartupMessages(library(tidyverse) )

ms <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/ms.csv") # load the data
# ls(as_tibble(ms))
```

# Guiding Question
## insert the guiding question here

# Overall Findings


## Conclusion  


# Recommendation



# Individual Findings
<!-- gist: plot variation in one of the spec variables -->
      <!-- plot covariation in the two -->


## Lexie Marinelli
Here is a plot of spec5 raw data and of the difference between spec5 and spec10 compared to spec5.
```{r, echo=FALSE}
#variation between spec5
ms %>%
  filter(spec5>0 & spec5< 10) %>%
  ggplot(ms, mapping=aes(x=spec5)) +geom_histogram(binwidth= 0.05, color= "blue") 
```
The first graph shows the relationship for spec5 as it gets larger, the occurance of those values faces an exponential decrease. 

```{r, echo=FALSE}
#covariation bt spec5 and spec10
ms %>%
  mutate(difference= spec5-spec10) %>%
  filter(spec5 > 0 & spec5< 25000 ) %>%
  ggplot(ms, mapping= aes(x=spec5, y=difference)) +geom_line()
```
The second graph tells us that as the value for spec5 increases, the difference between the values of spec5 and spec10 increase and oscillates to form a v shape. 


## Lindsay Gettel
```{r, echo=FALSE}

#raw spec10 data(for team)
ms%>%
  filter(spec10>0)%>%
  ggplot(mapping=aes(x=mass, y=spec10))+ geom_jitter()
ms%>%
  filter(spec10>0 & spec10<25)%>%
  ggplot(mapping=aes(x=spec10)) +geom_histogram(binwidth=0.5)

#histogram of difference between spec10 and spec5(for team)
ms%>%
  filter(spec5>0 & spec5<25, spec10>0 & spec10<25)%>%
  #summarize(spec5AVG=mean(spec5), spec10AVG=mean(spec10))%>%
  mutate(diff=spec10-spec5)%>%
  ggplot(mapping=aes(x=diff))+ geom_histogram(binwidth = 0.5)

#individual graphs, variation of spec10 and covaration
ms%>%
  filter(spec10>0 & spec10<30)%>%
  ggplot(mapping=aes(x=spec10))+ geom_histogram(binwidth=0.5)

ms%>%
  filter(spec10>0 & spec10<75)%>%
  ggplot(mapping=aes(x=mass, y=spec10))+ geom_jitter()+ geom_smooth(se=FALSE)+ ylim(0,50)
  
ms%>%
  filter(spec10>0 & spec10<50, spec5>0 & spec5<50)%>%
  ggplot()+ geom_density(mapping=aes(x=spec5), color="red")+ geom_density(mapping=aes(x=spec10), color="blue")#+ theme(legend=c("spec5", "spec10"), col=c("red", "blue"))
```




## Zhenlong Li


## Scott Baker
Here is a plot of the relationship between the spec5 and spec10 variables on a logarithmic scale 
```{r, echo=FALSE}
ggplot(data=ms, mapping=aes(x = mass, y = spec10)) + geom_point() + scale_x_log10() + scale_y_log10() + ggtitle(label = "Log Plot of spec10 vs. mass")

ggplot(data=ms, mapping=aes(x = mass, y = spec5)) + geom_point() + scale_x_log10() + scale_y_log10() + ggtitle(label = "Log Plot of spec5 vs. mass")

ggplot(data=ms, mapping=aes(x = spec5, y = spec10)) + geom_point() + scale_x_log10() + scale_y_log10() + ggtitle(label = "Log Plot of spec5 vs. spec10")

```
currently cant decide what shows the log data the best... will updated later

# Contributions
* Lindsay: 

* Lexie: 

* Li: 

* Scott: