---
title: "Lab_05"
author: "Fantastic Four"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#taken straight from Lab_5.html provided:
rm(list=ls()) # clears your environment variables
cat('\014')   # clears the console
suppressPackageStartupMessages(library(tidyverse) )

ms <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/ms.csv") # load the data
# ls(as_tibble(ms))
```

# Guiding Question
## insert the guiding question here

# Overall Findings


## Conclusion  


# Recommendation



# Individual Findings
<!-- gist: plot variation in one of the spec variables -->
      <!-- plot covariation in the two -->


## Lexie Marinelli
Here is a plot of spec5 raw data and of the difference between spec5 and spec10 compared to spec5.
```{r}
#variation between spec5
ms %>%
  filter(spec5>0 & spec5< 1.9) %>%
  ggplot(ms, mapping=aes(x=spec5)) +geom_histogram(binwidth= 0.05, color= "blue") 

#covariation bt spec5 and spec10
ms %>%
  mutate(difference= spec5-spec10) %>%
  filter(spec5 > 0 & spec5< 25000 ) %>%
  ggplot(ms, mapping= aes(x=spec5, y=difference)) +geom_line()
```


## Lindsay Gettel
```{r, echo=false}

#raw spec10 data(for team)
ms%>%
  filter(spec10>0)%>%
  ggplot(mapping=aes(x=mass, y=spec10))+ geom_jitter()

#ggplot(data=ms, mapping=aes(x=spec10)) +geom_histogram(binwidth=0.1)
"
#histogram of difference between spec10 and spec5(for team)
ms%>%
  filter(spec5>0, spec10>0)%>%
  summarize(spec5AVG=mean(spec5), spec10AVG=mean(spec10))%>%
  mutate(diff=spec10-spec5)%>%
  ggplot(mapping=aes(x=diff))+ geom_histogram(binwidth = 0.1)
  ggplot(mapping=aes(x=mass, y=diff))+ geom_jitter()"

#individual graphs, variation of spec10 and covaration
ms%>%
  filter(spec10>0)%>%
  ggplot()+ geom_histogram(mapping=aes(x=spec10),binwidth=0.1)
  ggplot(mapping=aes(x=mass, y=spec10))+ geom_jitter()+ geom_smooth(se=FALSE)
  
ms%>%
  filter(spec10>0 & spec5>0)%>%
  ggplot()+ geom_density(mapping=aes(x=spec5))+ geom_density(mapping=aes(x=spec10))
```




## Zhenlong Li


## Scott Baker
Here is a plot of the relationship between the spec5 and spec10 variables on a logarithmic scale 
```{r, echo=FALSE}
ggplot(data=ms, mapping=aes(x = mass, y = spec10)) + geom_point() + scale_x_log10() + scale_y_log10() + ggtitle(label = "Log Plot of spec10 vs. mass")

ggplot(data=ms, mapping=aes(x = mass, y = spec5)) + geom_point() + scale_x_log10() + scale_y_log10() + ggtitle(label = "Log Plot of spec5 vs. mass")

ggplot(data=ms, mapping=aes(x = spec5, y = spec10)) + geom_point() + scale_x_log10() + scale_y_log10() + ggtitle(label = "Log Plot of spec5 vs. spec10")

```
currently cant decide what shows the log data the best... will updated later

# Contributions
* Lindsay: 

* Lexie: 

* Li: 

* Scott: